\mychapter{Literature Review}

\vspace{0.5cm}

\subsection{Introduction}

Grading descriptive exam scripts is still one of the more labor-intensive jobs on any campus. Registration, scheduling, and result publication have mostly gone digital, but the actual marking of handwritten scripts hasn't caught up -- it's still mostly done by hand. To close that gap, researchers have looked at Optical Character Recognition (OCR), document image analysis, and web-based exam management systems as possible fixes \cite{article,inproceedings,8878791}.

Better scanning hardware, cloud computing, document image processing, and modern web frameworks have together made automated document handling, remote evaluation, and centralized exam management far more practical than they used to be. OCR and document layout analysis in particular have pushed paper digitization forward, turning scanned pages into something a computer can actually work with \cite{4376991,FRANCIS2025100532,docbank2020}. Even so, most systems in this space only tackle one piece of the puzzle -- none of them offer a full solution for descriptive admission script evaluation.

This chapter works through the existing literature on digital exam management systems, OCR, document layout analysis, PDF processing, digital script evaluation, and moderation. It ends by pointing out the gap that this thesis' proposed Digital Admission Script Evaluation and Moderation System (DASEMS) is built to fill.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Digital Examination Management Systems}

Centralized administration, secure logins, automated results, and better scheduling are where digital tools have made the biggest dent in exam management so far. Dandage et al. \cite{inproceedings} built an online exam and evaluation system that streamlines administration and result handling through a single web platform. Di Giusto \cite{article} found much the same thing from a different angle -- digital assessment cuts costs and speeds things up considerably, especially for objective-type exams.

Descriptive exams are a different story, though. Because grading a written answer takes human judgment, the process has stayed largely manual. Rajesh and Kanimozhi \cite{8878791} built a digitized paper-evaluation system using OCR and image processing to make document handling easier, but their work stops at digitizing the documents themselves -- it doesn't really touch question-wise evaluation, anonymous grading, or centralized moderation.

That leaves descriptive admission exams stuck with the same old problems: workload that's hard to distribute fairly, inconsistent grading across examiners, and administrative coordination that eats up time. All of which points to the need for a more complete digital evaluation framework.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Optical Character Recognition and Document Image Processing}

OCR is the backbone technology behind turning scanned pages into machine-readable text. Smith \cite{4376991} built the Tesseract OCR engine, and it's still one of the most widely used open-source options out there, thanks to its modular design and solid accuracy on printed text.

Deep learning has pushed OCR accuracy further still. Francis and Sangeetha \cite{FRANCIS2025100532} benchmarked several OCR models across different languages and mathematical notation, and found that deep-learning-based approaches beat traditional OCR by a clear margin on complex documents. Kumar et al. \cite{inbook} showed something complementary: pairing OpenCV-based preprocessing with OCR noticeably boosts recognition accuracy.

Document image processing picks up where OCR leaves off, analyzing page layout, spotting document structure, and pulling out the regions that actually matter. Li et al. put together the DocBank dataset \cite{docbank2020} specifically to support layout-analysis research, and separately, Zhou et al. \cite{Zhou2017EASTAE} and Baek et al. \cite{8953846} developed efficient scene-text-detection methods that can locate text regions even in cluttered documents.

For PDFs that already contain embedded text, tools like PyMuPDF and pdfplumber can pull that text out directly, coordinates and all, without needing OCR at all \cite{10.1145/3696474.3696731}. That approach doesn't help here, though -- handwritten admission scripts are scanned images with no embedded text layer, so there's nothing for these tools to extract.

Rather than lean fully on OCR, the system proposed in this thesis takes a template-based segmentation approach instead. Because every admission script follows the same fixed layout, question regions can be pulled out using predefined template coordinates once the document is aligned. OCR is kept in reserve as a tool for future, more intelligent document analysis, rather than being the core mechanism here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Digital Answer Script Evaluation and Moderation}

Digital evaluation platforms let teachers grade scanned scripts over the web, cutting out the need to physically pass scripts around. That opens the door to remote grading, centralized administration, and real-time progress tracking -- all of which trims down both time and administrative burden \cite{8878791,inproceedings}.

Many of these platforms also come with annotation tools, so examiners can highlight text, leave comments, and assign marks digitally, all without altering the original scanned document. Even with these tools in place, though, keeping grading consistent across multiple examiners is still hard, simply because descriptive assessment is subjective by nature.

One common fix in the literature is moderation: have multiple teachers independently grade the same script, then bring in a senior examiner to resolve any disagreement \cite{article}. This kind of setup tends to improve fairness and cut down on individual evaluator bias, all while keeping academic standards intact.

The DASEMS platform proposed here builds on these ideas by bringing anonymous evaluation through \textbf{s\_code}, automated teacher assignment, browser-based annotation, centralized moderation, and real-time evaluation tracking together into one system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Existing Technologies and Research Gap}

Progress in web technologies, OCR, document image processing, and cloud infrastructure has moved digital exam management forward quite a bit. Stacks built on React, Node.js, Express.js, MongoDB, Docker, and modern Python document-processing frameworks give developers a solid base for building evaluation platforms that can actually scale.

That said, the literature has some clear gaps. Most existing systems handle just one piece -- online exams, OCR-based digitization, or electronic marking -- in isolation. Very few cover the entire workflow of descriptive admission script evaluation: automatic question-wise segmentation, anonymous grading, automated teacher assignment, three-examiner moderation, centralized administration, and integrated result generation, all working together.

So there's still a real need for a single platform that can manage the whole lifecycle of descriptive admission script evaluation. DASEMS is built to fill exactly that gap, bringing together template-based document segmentation, role-based evaluation management, automated task assignment, browser-based annotation, moderation, and centralized result generation in one web platform.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Chapter Summary}

This chapter walked through existing research on digital exam management systems, OCR, document image processing, document layout analysis, and digital script evaluation. Taken together, the literature shows real progress on exam administration, document digitization, and electronic evaluation.

But it also makes clear that most systems address one component at a time rather than the full workflow. Features like anonymous evaluation, automatic question-wise segmentation, automated teacher assignment, centralized moderation, and integrated result generation are still underserved.

These gaps are what motivate DASEMS -- a system that pulls together modern web technologies, template-based document processing, and digital evaluation techniques to make university admission script evaluation more efficient, transparent, and reliable.