\mychapter{Proposed Methodology and System Design}
\vspace{0.5cm}

\subsection{Introduction}

This chapter walks through how the proposed \textbf{Digital Admission Script Evaluation and Moderation System (DASEMS)} was designed and built. It covers the system architecture, the design approach behind it, the technologies used, and how the major components talk to each other during the admission script evaluation process.

At its core, the system takes the conventional paper-based evaluation process and rebuilds it digitally -- role-based access control, secure document management, automated PDF processing, OCR-assisted question segmentation, browser-based evaluation, moderation, and centralized result generation, all in one platform. To keep evaluation fair, student identities stay hidden behind anonymous identifiers, and answer regions are pulled from uploaded scripts without ever touching the original document.

DASEMS is built as a set of separate modules: authentication, administration, document processing, evaluation, and moderation. The frontend runs on React with TypeScript, the backend on Node.js and Express.js, MongoDB handles the data, and Python takes care of PDF processing and OCR. Together, this gives a system that's scalable, easy to maintain, and well suited to the demands of descriptive admission script evaluation. The sections that follow go through the methodology and system design in more detail.

\subsection{Research Methodology}

Building \textbf{DASEMS} followed an incremental, modular approach to software development. Work started with identifying the problem clearly and reviewing the existing literature, which helped pin down where current admission evaluation systems fall short and where this research could add something new.

From there, the functional and non-functional requirements were worked out, which in turn shaped the system architecture, database design, and document processing workflow. Implementation then moved forward in stages -- each module was built, tested, and folded into the system one at a time rather than all at once.

The system breaks down into three main pieces: a React-based frontend, a Node.js backend API, and a Python service that handles PDF processing and question segmentation. Each of these was developed on its own before being brought together into a single platform.

Once implementation was done, functional and integration testing checked that each module worked correctly and that the evaluation workflow held together end to end. The finished system was then assessed against the research objectives, looking specifically at functionality, reliability, and usability.

Figure~\ref{fig:research_methodology} lays out the overall methodology followed in this study.

\begin{figure}[ht]
\centering

\begin{tikzpicture}[
>=Stealth,
process/.style={
draw,
rounded corners=5pt,
thick,
minimum width=7cm,
minimum height=0.8cm,
align=center,
font=\bfseries
},
arrow/.style={
->,
thick
}
]

\node[process,fill=blue!20]    (p1) at (0,0)      {Problem Identification};
\node[process,fill=green!20]   (p2) at (0,-1.5)   {Literature Review};
\node[process,fill=yellow!20]  (p3) at (0,-3.0)   {Requirement Analysis};
\node[process,fill=orange!20]  (p4) at (0,-4.5)   {System Design};
\node[process,fill=cyan!20]    (p5) at (0,-6.0)   {Implementation};
\node[process,fill=teal!20]    (p6) at (0,-7.5)   {System Integration};
\node[process,fill=pink!20]    (p7) at (0,-9.0)   {Testing and Validation};
\node[process,fill=gray!20]    (p8) at (0,-10.5)  {Performance Evaluation};

\draw[arrow] (p1)--(p2);
\draw[arrow] (p2)--(p3);
\draw[arrow] (p3)--(p4);
\draw[arrow] (p4)--(p5);
\draw[arrow] (p5)--(p6);
\draw[arrow] (p6)--(p7);
\draw[arrow] (p7)--(p8);

\end{tikzpicture}

\caption{Research methodology adopted for the development of DASEMS.}
\label{fig:research_methodology}

\end{figure}

\subsubsection{Development Approach}

Building the system incrementally kept both implementation and testing manageable. Rather than tackling everything at once, the system was split into independent modules that were built and integrated one after another, with each module checked over before moving on to the next. That kept complexity down and made the overall system more reliable.

Development moved through the following phases:

\begin{enumerate}

\item Requirement analysis and system planning.

\item Database design and backend API development.

\item Frontend implementation and user interface development.

\item PDF processing and question segmentation module development.

\item Evaluation workflow and teacher assignment implementation.

\item Moderation, result generation, and administrative dashboard development.

\item System integration, testing, and performance evaluation.

\end{enumerate}

This modular structure keeps the system easy to maintain and leaves room for future additions -- OCR-based handwriting recognition or AI-assisted evaluation, for instance -- without having to rework the underlying architecture.

\subsection{Requirement Analysis}

The requirements for \textbf{DASEMS} came out of looking closely at how admission scripts are currently evaluated and what each type of user is responsible for along the way. These requirements fall into two groups: functional and non-functional.

\subsubsection{Functional Requirements}

The system needs to support:

\begin{enumerate}
    \item Secure authentication with role-based access control.
    \item Management of admission sessions, subjects, question banks, teachers, and students.
    \item Upload and processing of scanned answer scripts with question-wise segmentation.
    \item Automatic assignment of evaluation tasks to subject teachers.
    \item Browser-based evaluation with digital annotation, comments, and mark submission.
    \item Moderation of conflicting evaluations by the Head Examiner.
    \item Result generation, report generation, and evaluation progress monitoring.
\end{enumerate}

\subsubsection{Non-Functional Requirements}

Beyond functionality, the system needs to hold up on these fronts:

\begin{itemize}
    \item \textbf{Security:} JWT authentication and role-based authorization.
    \item \textbf{Scalability:} Supports large-scale admission examinations.
    \item \textbf{Reliability:} Secure storage of documents and evaluation records.
    \item \textbf{Maintainability:} Modular architecture for future extensions.
    \item \textbf{Performance:} Efficient document processing and responsive user interface.
    \item \textbf{Availability:} Continuous access during the evaluation period.
\end{itemize}

\subsubsection{User Roles}

The system defines three user roles, each with a distinct set of responsibilities, summarized in Table~\ref{tab:user_roles}.

\begin{table}[ht]
\centering
\caption{User Roles and Responsibilities}
\label{tab:user_roles}
\begin{tabular}{|p{3.5cm}|p{10cm}|}
\hline
\textbf{User Role} & \textbf{Responsibilities} \\
\hline
Super Administrator &
Manage admission sessions, subjects, question banks, teachers, students, answer script uploads, and publish final results. \\
\hline
Head Examiner &
Assign evaluation tasks, monitor evaluation progress, moderate conflicting evaluations, and approve final marks. \\
\hline
Teacher &
Evaluate assigned questions, annotate answer scripts, assign marks, and submit evaluations. \\
\hline
\end{tabular}
\end{table}

\subsection{Overall System Architecture}

\textbf{DASEMS} is built as a modular, multi-layer system that keeps user interaction, business logic, document processing, and data management as separate, independent pieces. Splitting things up this way makes the system easier to scale, maintain, and integrate. The architecture breaks down into four layers:

\begin{enumerate}
    \item Presentation Layer
    \item Application Layer
    \item Document Processing Layer
    \item Data Layer
\end{enumerate}

Each layer handles a specific job, and together they support the full admission script evaluation workflow.

\subsubsection{Presentation Layer}

This layer is the web interface itself, built with React and TypeScript. It gives each type of user their own dashboard and talks to the backend through RESTful APIs.

\subsubsection{Application Layer}

Built on Node.js and Express.js, this layer handles authentication, user management, question banks, the evaluation workflow, and result generation. It's also the piece that coordinates traffic between the frontend, the database, and the document processing service.

\subsubsection{Document Processing Layer}

This runs as its own Python service, separate from the rest of the system. It takes in uploaded PDF answer scripts, runs template-based question segmentation, pulls out question-wise answer regions, and sends structured metadata back to the backend. OCR can be brought in as a supporting tool for scanned documents wherever needed.

\subsubsection{Data Layer}

MongoDB is the main database here, holding user information, student records, question banks, uploaded scripts, extracted answers, evaluation records, moderation data, and exam results.

\subsubsection{System Workflow}

Putting it all together, the workflow runs as follows:

\begin{enumerate}
    \item Upload admission answer scripts.
    \item Process PDF documents and extract question-wise answer regions.
    \item Automatically assign evaluation tasks to subject-specific teachers.
    \item Perform digital evaluation and submit marks.
    \item Moderate conflicting evaluations.
    \item Generate final examination results and reports.
\end{enumerate}

Because the layers are kept separate, each subsystem can be developed on its own, and the whole thing is easier to extend later on -- whether that's AI-assisted evaluation, more advanced OCR, or distributed document processing.

\begin{figure}[H]
\centering

\begin{tikzpicture}[
>=Stealth,
box/.style={
draw,
rounded corners=5pt,
thick,
minimum width=5cm,
minimum height=0.8cm,
align=center,
font=\bfseries
},
smallbox/.style={
draw,
rounded corners=5pt,
thick,
minimum width=2.3cm,
minimum height=0.7cm,
align=center,
font=\small
},
arrow/.style={->,thick}
]

%---------------- Top ----------------%

\node[box,fill=blue!20] (users) at (0,0) {Users};

\node[smallbox,fill=yellow!20] (admin) at (-3,-1.5) {Admin};

\node[smallbox,fill=yellow!20] (teacher) at (0,-1.5) {Teacher};

\node[smallbox,fill=yellow!20] (head) at (3,-1.5) {Head};

\node[box,fill=green!20] (frontend) at (0,-3.5)
{React Frontend};

\node[box,fill=cyan!20] (api) at (0,-5.2)
{REST API Communication};

\node[box,fill=orange!20] (backend) at (0,-7)
{Node.js + Express Backend};

%---------------- Bottom ----------------%

\node[box,fill=purple!20] (mongo) at (-4,-9.3)
{MongoDB Database};

\node[box,fill=red!20] (python) at (4,-9.3)
{Python Processor};

\node[smallbox,fill=purple!10] (m1) at (-4,-11)
{Student Data};

\node[smallbox,fill=purple!10] (m2) at (-4,-12.2)
{Question Bank};

\node[smallbox,fill=purple!10] (m3) at (-4,-13.4)
{Evaluations};

\node[smallbox,fill=purple!10] (m4) at (-4,-14.6)
{Audit Logs};

\node[smallbox,fill=red!10] (p1) at (4,-11)
{OCR Processing};

\node[smallbox,fill=red!10] (p2) at (4,-12.2)
{PDF Segmentation};

\node[smallbox,fill=red!10] (p3) at (4,-13.4)
{Answer Extraction};

\node[smallbox,fill=red!10] (p4) at (4,-14.6)
{Image Cropping};

%---------------- Connections ----------------%

\draw[arrow] (users)--(frontend);

\draw[arrow] (admin)--(frontend);
\draw[arrow] (teacher)--(frontend);
\draw[arrow] (head)--(frontend);

\draw[arrow] (frontend)--(api);
\draw[arrow] (api)--(backend);

\draw[arrow] (backend)--(mongo);
\draw[arrow] (backend)--(python);

\draw[arrow] (mongo)--(m1);
\draw[arrow] (m1)--(m2);
\draw[arrow] (m2)--(m3);
\draw[arrow] (m3)--(m4);

\draw[arrow] (python)--(p1);
\draw[arrow] (p1)--(p2);
\draw[arrow] (p2)--(p3);
\draw[arrow] (p3)--(p4);

\end{tikzpicture}

\caption{Overall System Architecture}
\label{fig:system_architecture}

\end{figure}

\subsection{System Modules}

\textbf{DASEMS} is organized into a set of independent functional modules, each handling one specific responsibility within the evaluation workflow. Keeping them separate this way makes the system more modular, easier to maintain, and easier to scale.

Figure~\ref{fig:overall_workflow} shows how the overall system workflow ties these modules together.

\begin{figure}[H]
\centering

\begin{tikzpicture}[
    node distance=0.9cm,
    >=Stealth,
    process/.style={
        rectangle,
        rounded corners=6pt,
        minimum width=7cm,
        minimum height=0.9cm,
        draw=black,
        thick,
        align=center,
        font=\bfseries
    },
    arrow/.style={
        ->,
        thick
    }
]

\node[process,fill=blue!20] (a)
{System Administration};

\node[process,fill=green!20,below=of a] (b)
{Admission Session Management};

\node[process,fill=yellow!25,below=of b] (c)
{Question Bank Management};

\node[process,fill=orange!25,below=of c] (d)
{Student Registration};

\node[process,fill=cyan!25,below=of d] (e)
{PDF Upload \& Processing};

\node[process,fill=purple!20,below=of e] (f)
{Answer Extraction};

\node[process,fill=red!20,below=of f] (g)
{Teacher Assignment};

\node[process,fill=pink!20,below=of g] (h)
{Teacher Evaluation};

\node[process,fill=teal!20,below=of h] (i)
{Moderation};

\node[process,fill=gray!25,below=of i] (j)
{Result Generation};

\draw[arrow] (a) -- (b);
\draw[arrow] (b) -- (c);
\draw[arrow] (c) -- (d);
\draw[arrow] (d) -- (e);
\draw[arrow] (e) -- (f);
\draw[arrow] (f) -- (g);
\draw[arrow] (g) -- (h);
\draw[arrow] (h) -- (i);
\draw[arrow] (i) -- (j);

\end{tikzpicture}

\caption{Overall Admission Examination Workflow}
\label{fig:overall_workflow}

\end{figure}

\begin{table}[H]
\centering
\caption{Major System Modules}
\label{tab:system_modules}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{|p{5cm}|p{9cm}|}
\hline
\textbf{Module} & \textbf{Description} \\
\hline

User Authentication \& Authorization &
Provides secure authentication and role-based access control using JWT authentication. \\
\hline

Admission Session \& Question Bank &
Manages admission sessions, departments, subjects, question banks, model answers, and marking guidelines. \\
\hline

Student Management \& Script Upload &
Maintains student information, generates anonymous \textbf{s\_code}, and supports secure upload of scanned answer scripts. \\
\hline

PDF Processing \& Question Segmentation &
Processes uploaded PDF scripts and extracts question-wise answer regions using template-based document segmentation. \\
\hline

Teacher Assignment &
Automatically assigns question-wise evaluation tasks balacing workload \\
\hline

Teacher Evaluation &
Provides browser-based evaluation with digital annotation, comments, and mark submission. \\
\hline

Moderation &
Compares multiple evaluations and forwards conflicting cases to the Head Examiner for final decision. \\
\hline

Result Generation \& Reporting &
Generates final examination results, evaluation reports, progress statistics, and audit logs. \\
\hline

\end{tabular}
\end{table}

\subsection{Database Design}

\textbf{DASEMS} runs on \textbf{MongoDB}, a NoSQL document-oriented database, to store all examination data. Its flexible, JSON-like document structure fits naturally with the kind of data this system handles -- admission sessions, question banks, student records, answer metadata, evaluations, and audit logs. Data is split across independent collections that link to one another through unique identifiers, which keeps things scalable, quick to query, and simpler to maintain over time.

\subsubsection{Database Collections}

Table~\ref{tab:database_collections} lists the main MongoDB collections used in the system.

\begin{table}[H]
\centering
\caption{Major Database Collections}
\label{tab:database_collections}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{|p{5cm}|p{9cm}|}
\hline
\textbf{Collection} & \textbf{Purpose} \\
\hline

User &
Stores user accounts, authentication credentials, roles, and profile information. \\
\hline

Admission Session &
Maintains admission sessions, departments, schedules, and session status. \\
\hline

Question Bank &
Stores questions, model answers, marking rubrics, and maximum marks. \\
\hline

Student &
Stores student information and anonymous \textbf{s\_code}. \\
\hline

Script &
Stores uploaded answer scripts and related metadata. \\
\hline

Answer &
Stores question-wise segmented answer regions and extraction metadata. \\
\hline

Evaluation Task &
Stores teacher assignments and evaluation status for each question. \\
\hline

Evaluation &
Stores marks, annotations, comments, and submitted evaluations. \\
\hline

Moderation &
Stores moderation decisions and final approved marks. \\
\hline

Audit Log &
Records important system activities and user actions. \\
\hline

\end{tabular}
\end{table}

\subsubsection{Entity Relationships}

Each admission session holds multiple students, subjects, and questions. Every student maps to a single uploaded answer script, which then gets broken down into several question-wise answers. Each of those extracted answers goes out to three teachers, so three independent evaluation records come back for it. If the marks from those three come in too far apart, the answer gets kicked up to the Head Examiner for review. Once evaluations are finalized, they're tied back to the corresponding student for result generation, and every step along the way gets logged in the audit trail.

\subsection{PDF Processing and Question Segmentation}

The PDF Processing and Question Segmentation module turns uploaded admission scripts into question-wise answer regions ready for digital evaluation. The original uploaded PDF is never touched -- instead, individual answer regions and their metadata are generated separately for evaluation.

The pipeline runs through PDF validation, page rendering, template alignment, question segmentation, answer region extraction, and metadata generation, with OCR available as a supporting technique for scanned documents when it's needed. Figure~\ref{fig:pdf_processing_workflow} lays out this workflow.

\subsubsection{PDF Upload and Validation}

Every uploaded script gets checked for file type and integrity first, then stored as a read-only document before any further processing happens.

\subsubsection{PDF Rendering}

Each page of the uploaded PDF is rendered into a high-resolution image using PyMuPDF, setting it up for the processing steps that follow.

\subsubsection{Template Alignment and Question Segmentation}

Rendered pages are lined up against a predefined answer sheet template. From there, question-wise answer regions are segmented and pulled out based on the template layout, without ever modifying the original document. OCR steps in as needed for document alignment or text detection.

\subsubsection{Question Detection}

Question detection runs as a two-stage, hybrid process. First, the processor tries to pull embedded PDF text and its positional coordinates directly. If there's no searchable text to work with, it falls back to OCR on the rendered page images to pick out printed question numbers and where they sit on the page. This keeps things fast when possible and still robust when it isn't.

\subsubsection{Answer Region Extraction}

Once question positions are known, the system works out the boundary of each answer region, crops it out as its own image, and generates metadata -- the anonymous student identifier, subject, question number, page number, coordinates, and image location. These records get stored for teacher assignment and evaluation, and the original PDF stays exactly as it was uploaded.

\subsubsection{Advantages of the Proposed Pipeline}

This pipeline leaves the original answer script untouched, automates the pulling out of question-wise answers, and takes a fair amount of administrative work off people's plates. Because the hybrid strategy only reaches for OCR when it's actually needed, it handles both searchable and scanned PDFs without wasting processing time. And since each extracted answer becomes its own self-contained unit, the system can assign teachers automatically, run evaluations in parallel, and moderate efficiently even at the scale of a large admission examination.

\subsubsection*{Algorithm 3.1: OCR-Based Answer Extraction}

\begin{verbatim}
Input:
    Uploaded PDF document

Output:
    Question-wise extracted answer images

Begin

Validate uploaded PDF

Store original PDF

For each page in the document do

    If searchable text exists then

        Extract text layer
        Detect question positions

    Else

        Render page image
        Perform OCR
        Detect question positions

    End If

    Identify answer boundaries

    Crop answer regions

    Save extracted answer image

    Generate metadata

End For

Store extracted answer records
Return extracted answers
End
\end{verbatim}

\begin{figure}[H]
\centering

\begin{tikzpicture}[
>=Stealth,
node distance=1cm,
box/.style={
rectangle,
rounded corners=6pt,
minimum width=6cm,
minimum height=0.9cm,
draw=black,
thick,
align=center,
font=\bfseries
},
arrow/.style={
->,
thick
}
]

% Nodes
\node[box,fill=blue!20] (upload) {PDF Upload};

\node[box,fill=green!20,below=of upload] (validation)
{File Validation};

\node[box,fill=yellow!25,below=of validation] (store)
{Store Original PDF};

\node[box,fill=orange!25,below=of store] (render)
{PDF Rendering};

\node[box,fill=cyan!25,below=of render] (align)
{Template Alignment};

\node[box,fill=purple!20,below=of align] (segment)
{Question Segmentation};

\node[box,fill=pink!20,below=of segment] (crop)
{Answer Region Cropping};

\node[box,fill=teal!20,below=of crop] (metadata)
{Metadata Generation};

\node[box,fill=gray!20,below=of metadata] (records)
{Create Answer Records};

% Connections
\draw[arrow] (upload) -- (validation);
\draw[arrow] (validation) -- (store);
\draw[arrow] (store) -- (render);
\draw[arrow] (render) -- (align);
\draw[arrow] (align) -- (segment);
\draw[arrow] (segment) -- (crop);
\draw[arrow] (crop) -- (metadata);
\draw[arrow] (metadata) -- (records);

\end{tikzpicture}

\caption{PDF Processing and Question Segmentation Workflow}
\label{fig:pdf_processing_workflow}

\end{figure}

\subsection{Teacher Assignment and Moderation}

Once questions are segmented, each extracted answer gets routed automatically to teachers in the relevant subject. The assignment logic weighs subject compatibility alongside current teacher workload, so tasks end up spread out fairly. Every answer goes to three separate teachers, and none of them can see how the others scored it, which keeps the assessment unbiased.

\subsubsection*{Algorithm 3.2: Teacher Assignment}

\begin{verbatim}
Input : Extracted Answer, Teacher List
Output: Assignment Records

Retrieve eligible teachers
Sort by workload
Assign three teachers
Store assignments

Return
\end{verbatim}

\subsubsection{Moderation}

Once all three evaluations for an answer come in, the system compares the marks. If they're close enough -- within a set threshold -- the final score gets calculated automatically. If not, the answer goes to the Head Examiner for a closer look.

The mark spread is calculated as

\begin{equation}
Spread=\max(M_1,M_2,M_3)-\min(M_1,M_2,M_3)
\end{equation}

If

\begin{equation}
Spread \leq T,
\end{equation}

the final mark is

\begin{equation}
FinalMark=\frac{M_1+M_2+M_3}{3}.
\end{equation}

Otherwise, the Head Examiner reviews the evaluations and determines the final mark.

\subsubsection*{Algorithm 3.3: Moderation}

\begin{verbatim}
Input : Three Evaluations
Output: Final Mark

Calculate mark spread

If spread <= threshold
    FinalMark = Average(M1,M2,M3)
Else
    Forward to Head Examiner
End If

Store final result

Return
\end{verbatim}

\subsection{Security and Authentication}

Authentication runs on encrypted passwords and JSON Web Tokens (JWT), and authorization is enforced based on each user's role.

Three roles are supported:

\begin{itemize}
    \item \textbf{Super Administrator:} Manages users, admission sessions, and examination data.
    \item \textbf{Head Examiner:} Oversees evaluation progress and moderates conflicting assessments.
    \item \textbf{Teacher:} Evaluates assigned questions and submits marks.
\end{itemize}

During evaluation, anonymous student identifiers (\textbf{s\_code}) keep assessment free of bias.

\subsection{Deployment Architecture}

DASEMS is deployed as a set of independent, modular components, which gives the system room to scale and more flexibility in how it's deployed.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    >=Stealth,
    node distance=1.4cm,
    box/.style={
        draw,
        rounded corners,
        thick,
        minimum width=4cm,
        minimum height=1cm,
        align=center,
        font=\bfseries
    },
    smallbox/.style={
        draw,
        rounded corners,
        thick,
        minimum width=3cm,
        minimum height=0.8cm,
        align=center,
        font=\small
    },
    arrow/.style={->,thick}
]

% Main Components
\node[box,fill=blue!20] (frontend) {React Frontend};

\node[box,fill=green!20,below=of frontend] (backend)
{Node.js API Server};

\node[box,fill=orange!20,below left=1.6cm and 2.5cm of backend] (db)
{MongoDB};

\node[box,fill=purple!20,below right=1.6cm and 2.5cm of backend] (python)
{Python Processor};

% Subcomponents
\node[smallbox,fill=orange!10,below=0.9cm of db] (uploads)
{Uploads\\Images};

\node[smallbox,fill=purple!10,below=0.9cm of python] (ocr)
{OCR\\Cropping};

% Connections
\draw[arrow] (frontend) -- (backend);

\draw[arrow] (backend) -- (db);
\draw[arrow] (backend) -- (python);

\draw[arrow] (db) -- (uploads);
\draw[arrow] (python) -- (ocr);

\end{tikzpicture}
\caption{Overall System Component Architecture}
\label{fig:system_component_architecture}
\end{figure}

The frontend runs on React and TypeScript, while the backend, built with Node.js and Express.js, handles authentication, evaluation, moderation, and the rest of the system's APIs.

PDF processing, OCR, question detection, and answer extraction all run through a separate Python service, kept independent of the web server. MongoDB holds user information, examination data, evaluations, moderation records, and final results. Every major service runs in its own Docker container, which keeps deployment simple and the system easy to move between environments.